### 2014-11-13 tex/operations.tex

In the field-access rewrite rule in~\figref{fig:fHeap}(a), after the
non-nullness check we extract the The type $C$ of the field is
extracted and invoke the summarize rewrite rule
in~\figref{fig:symInit} that performs the initialization of the field
if it points to uninitialized locations. Once the initialization is
complete, we create a new local reference $r^\prime$. An important
property of the references in the bi-partiate graph is that they are
\emph{immutable}. Hence we create a new local reference and assign it
to the initialized reference, and return the local reference. In order
to de-reference a field $r.f$ we define a helper function which is
called the value set.

\input{value-set-def}

In the post-condition of the rewrite rule we assign
the value set of input reference $r$ to the local reference $r^\prime$
and return the local reference $r^\prime$ in the next state.

Consider the graph shown in~\figref{fig:fHeap}(b) the reference
$\cfgnt{r}_2^i$ is created during the initialization of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. The
reference that is returned during the access of the field, however, is
$\cfgnt{r}_3^s$. The reference $\cfgnt{r}_3^s$ points to the value set of $\cfgnt{r}_2^i$
which are: $(\phi_{2a}\ \cfgnt{l}_\mathit{null})$, $(\phi_{2b}\ \cfgnt{l}_2)$, and
$(\phi_{2c}\ \cfgnt{l}_1)$. The values of the constraints $\phi_{2a}$,
$\phi_{2b}$, and $\phi_{2c}$ are defined in~\figref{fig:initHeap}(d).

The field-write rewrite rule in~\figref{fig:fHeap}(a) after the
non-nullness check we We look up the value of the base reference in
the environment $\eta(x)$.  The reference $\cfgnt{r}_x$ is the base pointer
whose field, $\cfgnt{r}_\mathit{curr}$ is being written to to while the
reference $r$ is the target reference. The set $\Psi_x$ contains
tuples $(\phi\ l\ \cfgnt{r}_\mathit{curr})$ of constraints, locations, and
references. These tuples represent access chains leading from $\cfgnt{r}_x$ to
the reference of the field, $\cfgnt{r}_\mathit{curr}$. The goal is to
overwrite the $\cfgnt{r}_\mathit{curr}$ references with the target
references. Since the target of the write is $r$, we first check that
the location constraint pairs of $\cfgnt{L}(r)$ are satisfiable when accessed
through the $\cfgnt{r}_x$ chain. This is accomplished by the strengthening
function.

\input{st-def}

Note that the write is conditional. In the case that $\phi$ is true
then $\cfgnt{r}_\mathit{curr}$ will point to the constraint location pairs of
$\cfgnt{L}(\cfgnt{r})$ while if $\phi$ is false then $\cfgnt{r}_\mathit{curr}$ will continue
to point to the constraint location pair it is currently pointing to.
Again since the references are immutable we create a new reference for
each $\cfgnt{r}_\mathit{curr}$ and point them to the target constraint
location pairs.

Consider the example shown in~\figref{fig:fHeap}(b) where we assign
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\ \cfgt{:=}\ \lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp\rp$. Note that
in~\figref{fig:initHeap}(c) $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ is represented by
$\cfgnt{r}_1^i$. After applying the field-write rule the reference $\cfgnt{r}_1^i$ is
replaced by the fresh reference $\cfgnt{r}_4$ which now points to target
 $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ represented by $\cfgnt{L}(\cfgnt{r}_2^i)$.

@@

The rewrite rules to compare two references in the symbolic summary
heap are shown in~\figref{fig:eqs}. The equals references-true rewrite
rule returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$ \emph{can} be
equal. In GSE semantics checking equality of references is a simple
comparision of two concrete object references. In the symbolic summary
heap, however, we compare sets of constraint location pairs pointed to
by each reference to determine if they could be equal.

Consider the equals reference-true rewrite rule
in~\figref{fig:eqs}. In order to check equality we construct three
sets of constraints. For all constraint location pairs $(\phi_0\ \cfgnt{l}_0)
\in \cfgnt{L}(\cfgnt{r}_0)$ and $(\phi_1\ \cfgnt{l}_1) \in \cfgnt{L}(\cfgnt{r}_1)$ such that $\cfgnt{l}_0$ and $\cfgnt{l}_1$
are the same ($\cfgnt{l}_0 = \cfgnt{l}_1$) we create conjunctions of the constraints
$\phi_0 \wedge \phi_1$ and add them to set $\Phi_\alpha$. Intutively,
$\Phi_\alpha$ contains all constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may
point to the same location in the symbolic summary heap. The second
set, $\Phi_0$, contains constraints under which the reference $\cfgnt{r}_0$
points corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and $\cfgnt{r}_0$ \emph{does
  not} point to those locations under any constraint.

We use the three sets of constraints to update the current global heap
constraint $\phi_g$ and create a new global heap constraint
$\phi_g^\prime$. We add to $\phi_g$ the disjunction of the constraints
in $\Phi_\alpha$ to indicate that if any of the constraints are
satisfiable, then references $\cfgnt{r}_0$ and $\cfgnt{r}_1$ can be
equal. Furthermore, we had to the global heap constraint,
$\phi_g^\prime$, the conjunctions of negations to the constraints in
$\Phi_0$ and $\Phi_1$. This indicates for locations that are not
common to the references, the negations of their constraints are
satisfiable. Before the rewrite rule returns true we check the
satisfiability of the updated global heap constraint.

Consider the example in~\figref{fig:initHeap}(c). In order to to check
if $\cfgnt{r}_1^i$ and $\cfgnt{r}_2^i$ are equal we first get the constraint location
pairs associated with each of the references:
\[
\cfgnt{L}(\cfgnt{r}_1^i) = \{ (\phi_{1a}\; \cfgnt{l}_\mathit{null})\; (\phi_{1b}\; \cfgnt{l}_1) \} 
\]
\[
\cfgnt{L}(\cfgnt{r}_2^i) = \{ (\phi_{2a}\; \cfgnt{l}_\mathit{null})\; (\phi_{2b}\; \cfgnt{l}_2)\; (\phi_{2c}\; \cfgnt{l}_1) \} \\
\]
\noindent{The three constraint sets are:} 
\[
\Phi_\alpha = \{ (\phi_{1a}\; \wedge\; \phi_{2a} ) (\phi_{1b}\; \wedge\; \phi_{2c} ) \}\;
\Phi_0 = \{ \}\; \Phi_1 = \{ \phi_{2b}\} \\
\]
\noindent{Finally the global constraint is} 
\[
\phi^\prime = \mathit{true} \wedge [ (\phi_{1a}\; \wedge\; \phi_{2a} )\vee (\phi_{1b}\; \wedge\; \phi_{2c} ) ] \wedge \neg\phi_{2b} 
\]
\noindent{The equals references-false is the logical dual of the
  reference-true rewrite rule. The references-false is shown
  in~\figref{fig:eqs}.}

>>>>

In the field-access rewrite rule in~\figref{fig:fHeap}(a), after the
non-nullness check, the rules gets the field type $\cfgnt{C}$ and uses
the summarize machine,\figref{fig:symInit}, to get a new summary heap
that is initialized on the field $\cfgnt{f}$. The summary heap after
initialization is further modified in the final state of the field
access to include a new stack reference pointing to the value set from
the read. The new stack reference is the return value from the operation.
\input{value-set-def}
The value set computation is local: it only considers
constraint locations on the reference $\cfgnt{r}$. Further, it only
includes things that are feasible under the current path condition
$\phi_g$. It also ensures that the access path valid.

There is a constraint from the base reference and a given
location. Once a location is selected, the field is accessed to get
another reference, and from that reference there is a set of locations
each with a constraint. The first constraint from the base reference
plus this second constraint to the actual location of the field
comprise the access path. For each member of the value set, these
constraints together with the global path constraint must be satisfied
on some aliasing assignment.

\begin{comment}invoke the summarize rewrite rule
in~\figref{fig:symInit} that performs the initialization of the field
if it points to uninitialized locations. Once the initialization is
complete, we create a new local reference $r^\prime$. An important
property of the references in the bi-partiate graph is that they are
\emph{immutable}. Hence we create a new local reference and assign it
to the initialized reference, and return the local reference. In order
to de-reference a field $r.f$ we define a helper function which is
called the value set.

In the post-condition of the rewrite rule we assign
the value set of input reference $r$ to the local reference $r^\prime$
and return the local reference $r^\prime$ in the next state.
\end{comment}

Consider the graph shown in~\figref{fig:fHeap}(b). The goal is to
complete the operation $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$. The
machine first evaluates the local variable look up to get the base
reference: $\eta(\cfgt{this}) = \cfgnt{r}^s_0$. From the base
reference, field access is able to complete returning the stack
reference $\cfgnt{r}_3^s$ in the figure. The reference $\cfgnt{r}_3^s$
points to the value set of $\cfgnt{r}_2^i$:
$(\phi_{2a}\ \cfgnt{l}_\mathit{null})$, $(\phi_{2b}\ \cfgnt{l}_2)$,
and $(\phi_{2c}\ \cfgnt{l}_1)$.

For field-write in~\figref{fig:fHeap}(a), after the
non-null check and strengthening of the global heap constraint, it computes the current references associated with the field in every location in $\Psi_\cfgnt{x}$. Please note that the reference $\cfgnt{r}_x$ is the base reference
whose field, $\cfgnt{r}_\mathit{cur}$ is being written to while the
reference $r$ is the target reference. The set $\Psi_x$ contains
tuples $(\phi\ l\ \cfgnt{r}_\mathit{cur})$ of constraints, locations, and
references. These tuples represent access chains leading from $\cfgnt{r}_x$ to
the reference of the field, $\cfgnt{r}_\mathit{cur}$. The goal is to change the fields to no longer point to $\cfgnt{r}_\mathit{cur}$, but rather fresh references that have both the original locations before the write and the locations from the write in the value sets (i.e., conditional write). Since the target of the write is $r$, the rule checks that
the location constraint pairs of $\cfgnt{L}(\cfgnt{r})$ are satisfiable when accessed
through the $\cfgnt{r}_x$ chain in the strengthening function.
\input{st-def}
Constraints on locations are strengthened with new aliasing conditions
and those that are feasible with the current path constraint are
retained.

Strengthening in the field write creates a value set the
contains feasible locations if the write happens on that location (the
first call) and feasible locations if the write does not happen on
that location (the second call). This is the conditional in the write. In the case that $\phi$ is true
then $\cfgnt{r}_\mathit{cur}$ will point to the constraint location pairs of
$\cfgnt{L}(\cfgnt{r})$ while if $\phi$ is false then $\cfgnt{r}_\mathit{cur}$ will continue
to point to the constraint location pair it is currently referenced.
As the references are immutable, the rule creates fresh reference for
each $\cfgnt{r}_\mathit{cur}$ and points them to the appropriate value sets.

Consider again the example shown in~\figref{fig:fHeap}(b) where we assign
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\ \cfgt{:=}\ \lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp\rp$. Note that
in~\figref{fig:initHeap}(c) $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ is represented by
$\cfgnt{r}_1^i$. After applying the field-write rule the reference $\cfgnt{r}_1^i$ is
replaced by the fresh reference $\cfgnt{r}_4$ which now points to target
 $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ represented by $\cfgnt{L}(\cfgnt{r}_2^i)$.

@@

The rewrite rules to compare two references in the symbolic summary
heap are shown in~\figref{fig:eqs}. The equals references-true rewrite
rule returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$ \emph{can} be
equal. In GSE semantics, checking equality of references is a simple
comparision of two concrete object references. In the symbolic summary
heap, however, we compare sets of constraint location pairs pointed to
by each reference to determine if equality is possible.

Consider the equals reference-true rewrite rule
in~\figref{fig:eqs}. In order to check equality, three
sets of constraints are used. For all constraint location pairs $(\phi_0\ \cfgnt{l}_0)
\in \cfgnt{L}(\cfgnt{r}_0)$ and $(\phi_1\ \cfgnt{l}_1) \in \cfgnt{L}(\cfgnt{r}_1)$ such that $\cfgnt{l}_0$ and $\cfgnt{l}_1$
are the same ($\cfgnt{l}_0 = \cfgnt{l}_1$), the rule creates conjunctions of the constraints
$\phi_0 \wedge \phi_1$ and add them to set $\Phi_\alpha$. Intutively,
$\Phi_\alpha$ contains all constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may
point to the same location in the symbolic summary heap. The second
set, $\Phi_0$, contains constraints under which the reference $\cfgnt{r}_0$
points to corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and $\cfgnt{r}_0$ \emph{does
  not} point to those locations under any constraint.

We use the three sets of constraints to update the current global heap
constraint $\phi_g$ and create a new global heap constraint
$\phi_g^\prime$. We add to $\phi_g$ the disjunction of the constraints
in $\Phi_\alpha$ to indicate that if any of the constraints are
satisfiable, then references $\cfgnt{r}_0$ and $\cfgnt{r}_1$ can be
equal. Furthermore, we had to the global heap constraint,
$\phi_g^\prime$, the conjunctions of negations to the constraints in
$\Phi_0$ and $\Phi_1$. This indicates for locations that are not
common to the references, the negations of their constraints are
satisfiable. The rule does not complete (i.e., is not feasible) if the new global constraint is not satisfied on any aliasing assignment. In such a case, the true outcome is not possible on any heap in the summary.

Before the rewrite rule returns true we check the
satisfiability of the updated global heap constraint.

Consider the example in~\figref{fig:initHeap}(c). In order to to check
if $\cfgnt{r}_1^i$ and $\cfgnt{r}_2^i$ are equal we first get the constraint location
pairs associated with each of the references:
\[
\cfgnt{L}(\cfgnt{r}_1^i) = \{ (\phi_{1a}\; \cfgnt{l}_\mathit{null})\; (\phi_{1b}\; \cfgnt{l}_1) \} 
\]
\[
\cfgnt{L}(\cfgnt{r}_2^i) = \{ (\phi_{2a}\; \cfgnt{l}_\mathit{null})\; (\phi_{2b}\; \cfgnt{l}_2)\; (\phi_{2c}\; \cfgnt{l}_1) \} \\
\]
\noindent{The three constraint sets are:} 
\[
\Phi_\alpha = \{ (\phi_{1a}\; \wedge\; \phi_{2a} ) (\phi_{1b}\; \wedge\; \phi_{2c} ) \}\;
\Phi_0 = \{ \}\; \Phi_1 = \{ \phi_{2b}\} \\
\]
\noindent{Finally the global constraint is} 
\[
\phi^\prime = \mathit{true} \wedge [ (\phi_{1a}\; \wedge\; \phi_{2a} )\vee (\phi_{1b}\; \wedge\; \phi_{2c} ) ] \wedge \neg\phi_{2b} 
\]
\noindent{The equals references-false is the logical dual of the
  reference-true rewrite rule. The references-false is also shown
  in~\figref{fig:eqs}.}
